---
title: "WindCurves: A Tool to Fit Wind Turbine Power Curves"
author: "Neeraj Bokde (neerajdhanraj@gmail.com) and Andres Feijoo (afeijoo@uvigo.es)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WindCurves: A Tool to Fit Wind Turbine Power Curves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction:

This is a Vignettes of R package, `WindCurves`. The package `WindCurves` is a tool used to fit the wind turbine power curves. It can be useful for researchers, data analysts/scientist, practitioners, statistians and students working on wind turbine power curves. The salient features of `WindCurves` package are:

- Fit the power curve with Weibull CDF, Logistic [1] and user defined techniques.
- Comparison and visualization of the performance of curve fitting techniques.
- Utilise as a Testbench to compare performace of user defined curve fitting techniques with user defined error measures.
- Availability of Dataset [2] on the power curves of wind turbine from four major manufacturers: Siemens, Vestas, REpower and Nordex.
- Feature to extract/capture Speed Vs Power discrete points from power curve image

## Instructions to Use:

- A power curve can be fitted with the `WindCurves` package from a discrete samples of wind turbine power curves provided by turbine manufacurers as:

```{r, fig.show='hold', fig.height=5, fig.width=7}
library(WindCurves)
data(pcurves)
s <- pcurves$Speed
p <- pcurves$`Nordex N90`
da <- data.frame(s,p)
x <- fitcurve(da)
x
validate.curve(x)
plot(x)
```




- User can utilize `WindCurves` package as a testbench so that a new curve fitting technique can be compared with above mentioned existing environment.  

- - The user defined function (similar to `random()`) should be saved in .R format and it should return the fitted values in terms of "Power values". The `random()` function used in this example is stored as .R file with `dump()` function and made available in the `WindCurves` package and written as:

```{r, fig.show='hold', fig.height=5, fig.width=7}
random <- function(x)
{
  data_y <- sort(sample(1:1500, size = 25, replace = TRUE))
  d <- data.frame(data_y)
  return(d)
}
dump("random")
rm(random)
```

- A `random()` function is attached in the comparison the parameters `MethodPath` and `MethodName` as shown below, where `MethodPath` is a location of the function proposed by the user and `MethodName` is a name given to the function.

```{r, fig.show='hold', fig.height=5, fig.width=7}
library(WindCurves)
data(pcurves)
s <- pcurves$Speed
p <- pcurves$`Nordex N90`
da <- data.frame(s,p)
x <- fitcurve(data = da, MethodPath = "source('dumpdata.R')", MethodName = "Random values")

## The user can specify .R files from other locations as:
# x <- fitcurve(data = da, MethodPath = "source('~/WindCurves/R/random.R')", MethodName = "Random values")
validate.curve(x)
plot(x)
```

- Also, `WindCurves` allows user to add more error measures with `validate.curve()` function as explained below:

Consider `error()` is a function which uses two vectors as input and returns a error value with a specific error measure, such as RMSE or MAPE as shown below:

```{r, fig.show='hold', fig.height=5, fig.width=7}
# PCV as an error metric
error <- function(a,b)
{
d <- (var(a) - var(b)) * 100/ var(a)
d <- as.numeric(d)
return(d)
}
dump("error")
rm(error)
```

The effect of this function can be seen in the results obtained with `Validate.curve()` function as:

```{r, fig.show='hold', fig.height=5, fig.width=7}
library(WindCurves)
data(pcurves)
s <- pcurves$Speed
p <- pcurves$`Nordex N90`
da <- data.frame(s,p)
x <- fitcurve(da)
validate.curve(x = x, MethodPath = "source('dumpdata.R')", MethodName = "New Error")
plot(x)
```

Similarly, user can compare various techniques used for wind turbine power curve fitting.

- Dataset [2] on the power curves of wind turbine from four major manufacturers: Siemens, Vestas, REpower and Nordex. These datsset represent wind turbine power output in 'kW' against wind speed in 'metres per second' as shown below:

```{r, fig.show='hold', fig.height=5, fig.width=9}
data(pcurves)
pcurves
```

- Further, `WindCurves` package provides feature to extract/capture Speed Vs Power discrete points from power curve image. This can be achieved with the following simple instruction:

```{r, fig.show='hold', fig.height=5, fig.width=7}
#img2points("image.jpeg")
```

where, `image.jpeg` is the name of power curve image from which discrete points are to extracted. The procedure of extraction is as follows:

- With the command `img2points("image.jpeg")`, an image will get appeared in the Viewer.
- Click on leftmost point of the X-axis of the curve image and assign the respective appropriate value.
- click on rightmost point on X-axix, downmost point of Y-axis and topmost point of Y-axis, and assign their values, respectively.
- click of the points on the curve image. By default, 15 points can be captured from the curve image, but user can set the desired number of points with `n` parameter in `img2points()` function.
- This function returns a data.frame with two columns, i.e., wind speed and wind power, which can be used as input data for `fitcurve()` function.





## References

> [1] D. Villanueva and A. E. Feij´oo, “Reformulation of parameters of the logistic function applied to power curves of wind turbines,” Electric Power Systems Research, vol. 137, pp. 51–58, 2016.([via](https://www.sciencedirect.com/science/article/pii/S0378779616300992))

> [2] Iain Staffell, “Wind turbine power curves, 2012” ([via](https://goo.gl/tD2JW6))
